---
title: "nenswniche : a package for rapidly estimating multiple taxa ranges and habitat suitability"
authors: "Hugh Burley, Shawn Laffan, Will Cornwell, Adrian Fisher"
date: "October 2021"
output:
  github_document:
  toc: true   
toc_depth: 4            
toc_float: true
number_sections: false  
vignette: >
%\VignetteIndexEntry{README}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
  keep_md: true
  theme: united     
highlight: tango        
css: styles.css
revealjs::revealjs_presentation:
  dev: 'svg'
chunk_output_type: console
self_contained: false
reveal_plugins: ["notes", "search"]
reveal_options:
  slideNumber: true
previewLinks: true
word_document:
  always_allow_html: yes
---

\

The text and code below summarises a workflow in R that can be used to relatively rapidly assess the environmental range of a taxa within Australia, from downloading occurrence records, through to creating maps of predicted climatic suitability across Australia at 1km*1km resolution. An example of this work is published in the journal Science of the Total Environment ::

\

Burley, H., Beaumont, L.J., Ossola, A., et al. (2019) Substantial declines in urban tree habitat predicted 
under climate change. Science of The Total Environment, 685, 451-462.

https://www.sciencedirect.com/science/article/pii/S0048969719323289#f0030 

\

To install, run :

```{r eval=FALSE, echo=TRUE, message=TRUE, warning=FALSE}

## Set env
rm(list = ls())
#if (!Sys.getenv("JAVA_TOOL_OPTIONS")) {
if (all(Sys.getenv("JAVA_HOME")=="")) {
  Sys.setenv(JAVA_HOME='C:\\Program Files\\Java\\jre1.8.0_321')
}

if (all(Sys.getenv("JAVA_TOOL_OPTIONS")=="")) {
  options(java.parameters = "-Xmx64G")
}

options(warn=0)

## Function to load or install packages
ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg))
    install.packages(new.pkg, dependencies = TRUE, repos = "https://cran.csiro.au/")
  sapply(pkg, require, character.only = TRUE)
}


## Load packages
#devtools::install_github("HMB3/nenswniche")
library(nenswniche)
data('sdmgen_packages')
ipak(sdmgen_packages)


## The functions expect these folders,
main_dir             <- paste0(getwd(), "/")
tempdir              <- './TEMP/'
ALA_dir              <- './data/ALA/'
INV_dir              <- './data/ALA/Insects/'
check_dir            <- './data/ALA/Insects/check_plots/'
out_dir              <- './output/'

inv_rs_dir           <- './output/invert_maxent_raster_update/'
inv_back_dir         <- './output/invert_maxent_raster_update/back_sel_models/'
inv_full_dir         <- './output/invert_maxent_raster_update/full_models/'
inv_results_dir      <- './output/invert_maxent_raster_update/results/'

plant_rs_dir         <- './output/plant_maxent_raster_update/'
plant_back_dir       <- './output/plant_maxent_raster_update/back_sel_models/'
plant_full_dir       <- './output/plant_maxent_raster_update/full_models/'
plant_results_dir    <- './output/plant_maxent_raster_update/results/'

veg_dir              <- './data/Remote_sensing/Veg_data/Forest_cover/'
inv_habitat_dir      <- './output/invert_maxent_raster_update/Habitat_suitability/'
inv_records_dir      <- './output/invert_maxent_raster_update/Habitat_suitability/SDM_point_data/'
inv_inters_dir       <- './output/invert_maxent_raster_update/Habitat_suitability/SDM_Veg_intersect/'
inv_thresh_dir       <- './output/invert_maxent_raster_update/Habitat_suitability/SDM_thresholds/'
inv_fire_dir         <- './output/invert_maxent_raster_update/Habitat_suitability/FESM_SDM_intersect/'

plant_habitat_dir    <- './output/plant_maxent_raster_update/Habitat_suitability/'
plant_inters_dir     <- './output/plant_maxent_raster_update/Habitat_suitability/Veg_intersect/'
plant_thresh_dir     <- './output/plant_maxent_raster_update/Habitat_suitability/SDM_thresholds/'
plant_fire_dir       <- './output/plant_maxent_raster_update/Habitat_suitability/FESM_SDM_intersect/'


dir_list <- c(tempdir, ALA_dir, 
              INV_dir, check_dir, out_dir, inv_rs_dir, inv_back_dir, inv_full_dir, inv_results_dir,
              plant_rs_dir, plant_back_dir, plant_full_dir, plant_results_dir, veg_dir,
              inv_habitat_dir, inv_inters_dir, inv_thresh_dir, inv_fire_dir,
              plant_habitat_dir, plant_inters_dir, plant_thresh_dir, plant_fire_dir)


## Create the folders if they don't exist.
for(dir in dir_list) {
  
  if(!dir.exists(paste0(main_dir, dir))) {
    message('Creating ', dir, ' directory')
    dir.create(file.path(main_dir, dir), recursive = TRUE) 
    
  } else {
    message(dir, ' directory already exists')}
}


## Try and set the raster temp directory to a location not on the partition, to save space
rasterOptions(memfrac = 0.9,
              tmpdir  = tempdir)

terraOptions(memfrac = 0.9, 
             tempdir = tempdir) 

```

\
\
\

# Background

This code is being developed at UNSW, as part of a project investigating
the impacts of the 2019/2020 bush fires on Insects in the North East Forests 
of New South Wales. The aim is to create a pipeline that rapidly assesses 
the habitat suitability of the threatened insect taxa under current 
environmental conditions. 

\

# STEP 1 :: Get spp lists

\

The backbone of the R workflow is a list of (taxonomically Ridgey-Didge!) Taxa names 
that we supply. The analysis is designed to process data for one taxa at a time, 
allowing taxa results to be updated as required. Unfortunately, Australia's insects
are not very well sampled...so we can analyse at the family level.

\

Let's use all the insect families that we think might be threatened by the 2019/2020 fires. 

\

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=TRUE}


## To Do :
## 1). Check the errors that Finlay may have found
## 2). Clean up the folders


## get target taxa
data('insect_data_families')
data('all.insect.families')
data('all.insect.genera')
data('all.insect.spp')

data('target.insect.spp')
data('target.insect.genera')
data('target.insect.families')
data('target.host.plants')
data('all.insect.plant.spp')


## Full list of analysis taxa
## Full list of analysis taxa, 
analysis_taxa   <- str_trim(c(target.insect.spp, 
                              target.insect.genera, 
                              target.insect.families)) %>% unique()

taxa_qc <- read_excel(paste0(inv_habitat_dir, 'SDM_point_data/SDM_target_species.xlsx'),
                      sheet = 'Invert_QA_check')

taxa_remain <- taxa_qc %>% 
  filter(QC == "X") %>%
  filter(is.na(Morpho)) %>%
  dplyr::select(Binomial) %>% 
  .$Binomial %>% c(., 'Rayieria basifer') %>% sort()


```


The taxa list is supplied to a series of functions to calculate environmental ranges and habitat 
suitability. The initial functions download all taxa records from the Atlas and living Australia
(https://www.ala.org.au/) and the Global Biodiversity Information Facility (GBIF, https://www.gbif.org/). 
The taxa data are downloaded as individual .Rdata files to the specified folders, which must exist first, 
without returning anything.

\


# STEP 2 :: Combine taxa occurrence data

\

This pipeline combines taxa occurrence points with raster data. It was developed using worldclim 
climate raster data, but it can take any set of rasters. All the raster data is available here :: 

\

https://drive.google.com/open?id=1T5ET5MUX3-lkqiN5nNL3SZZagoJlEOal. 

\

Let's create all the raster data we need. First let's get the rasters needed for SDM modelling :
- Climate
- Vegetation
- Topography
- Soil

\

\

# 250m Climate, Soil and terrain variables

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=FALSE}

aus.climate.veg.grids.250m <- raster::stack(
  list.files('./data/CSIRO_layers/250m/AUS/',      pattern =".tif", full.names = TRUE))

east.climate.veg.grids.250m  <- raster::stack(
  list.files('./data/CSIRO_layers/250m/FESM_EXT/', pattern =".tif", full.names = TRUE))


names(aus.climate.veg.grids.250m)[1:11] <- 
  names(east.climate.veg.grids.250m)[1:11] <- 
  c("Tree_canopy_peak_foliage_total",
    "Plant_cover_fraction_0_5m", 
    "Plant_cover_fraction_5_10m",  
    "Plant_cover_fraction_10_30m",      
    "Plant_cover_fraction_30m",
    "Total_Plant_cover_fraction",  
    "Tree_canopy_height_25th", 
    "Tree_canopy_height_50th", 
    "Tree_canopy_height_75th",   
    "Tree_canopy_height_95th",   
    "Tree_canopy_peak_foliage")

identical(names(east.climate.veg.grids.250m),
          names(aus.climate.veg.grids.250m))

## Now remove the individual rasters
rm(list = ls(pattern = 'aus_'))
rm(list = ls(pattern = 'east_'))

aus_annual_precip       <- raster('./data/CSIRO_layers/250m/AUS/Extra/Annual_precip_WGS84.tif')
aus_annual_precip_alb   <- raster('./data/CSIRO_layers/250m/AUS/Extra/Annual_precip_GDA_ALB.tif')


## Should be 1km*1km, It should havle a value of 1 for land, and NA for the ocean
aus_annual_precip_alb[aus_annual_precip_alb > 0] <- 1
template_raster_250m <- aus_annual_precip_alb
gc()

```


\

Next we filter the records to those taken after 1950, and those inside the raster boundaries 
(i.e. taxa records in the ocean according to the raster boundaries will be excluded).

\

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=FALSE}

## Combine ALA data, and filter to records on land taken > 1950
## The climate data is the worldclim version 2.0
## Also, add a switch for the 1km filter - we don't want to do this anymore, too few records
data('ALA_keep')


## Take a data dump from the ALA of all the reptiles, 
## and format it to the columns names we want for the niche analyses
Insects_ALA      <- readRDS('./data/ALA/Insects/INVERTS_2022_1205.rds')



ALA.LAND.INV.SPP <- format_ala_dump(ALA_table     = head(Insects_ALA, 1000),
                                    record_type   = "ALA",
                                    keep_cols     = ALA_keep,
                                    year_filt     = FALSE,
                                    unique_cells  = FALSE,
                                    world_raster  = aus_annual_precip)


ALA.LAND.INV.GENUS <- ALA.LAND.INV.SPP %>% dplyr::select(-searchTaxon) %>% 
  dplyr::rename(searchTaxon = genus)   %>% dplyr::select(searchTaxon, everything())


ALA.LAND.INV.FAM   <- ALA.LAND.INV.SPP %>% dplyr::select(-searchTaxon) %>% 
  dplyr::rename(searchTaxon = family)  %>% dplyr::select(searchTaxon, everything())


'Rayieria basifer'            %in% ALA.LAND.INV.SPP$searchTaxon
'Amphistomus cunninghamensis' %in% ALA.LAND.INV.SPP$searchTaxon
'Eritingis trivirgata'        %in% ALA.LAND.INV.SPP$searchTaxon
'Asteron grayi'               %in% ALA.LAND.INV.SPP$scientificName


ALA.LAND.INV.SPP.SF <- SpatialPointsDataFrame(coords      = ALA.LAND.INV.SPP %>% 
                                                dplyr::select(lon, lat) %>% as.matrix(),
                                              data        = ALA.LAND.INV.SPP,
                                              proj4string = CRS("+init=epsg:4326")) %>%
  
  st_as_sf() %>% 
  st_transform(., st_crs(3577))


poly <- st_read('data/Spatial_data/Study_areas/AUS_2016_AUST.shp') %>% 
  st_transform(., st_crs(3577)) %>% as_Spatial()


## Check a few of the taxa now
'Rayieria basifer'            %in% ALA.LAND.INV.SPP.SF$searchTaxon
'Amphistomus cunninghamensis' %in% ALA.LAND.INV.SPP.SF$searchTaxon
'Eritingis trivirgata'        %in% ALA.LAND.INV.SPP.SF$searchTaxon
'Asteron grayi'               %in% ALA.LAND.INV.SPP.SF$scientificName


Amphistomus.cunninghamensis <- ALA.LAND.INV.SPP.SF %>% 
  filter(searchTaxon == 'Amphistomus cunninghamensis')


## For all the spatially incorrect taxa, plot them to see what's happening
for(taxa in taxa_remain) {
  
  ## taxa = target.insect.spp[79] 
  if(taxa %in% unique(ALA.LAND.INV.SPP.SF$searchTaxon)) {
    
    map <- paste0(inv_records_dir, taxa, '_ALA_POINTS.gpkg')
    
    if(!file.exists(map)) {
      
      message('Subsetting ', taxa, ' geo-package')
      save_name <- gsub(' ', '_', taxa)
      taxa_occ <- ALA.LAND.INV.SPP.SF %>% .[.$searchTaxon %in% taxa, ] 
      
      st_write(taxa_occ %>% st_as_sf(), 
               dsn    = paste0(inv_records_dir, taxa, '_ALA_POINTS.gpkg'), 
               layer  = paste0(taxa, '_SDM_points'), 
               quiet  = TRUE,
               append = FALSE)
      
      png(paste0(inv_records_dir, save_name, '_ALA_POINTS.png'),
          11, 4, units = 'in', res = 400)
      
      plot(st_geometry(taxa_occ), legend = FALSE, col = 'red')
      plot(poly, add = TRUE)
      dev.off()
      
      
    } else {
      message(taxa, ' already checked')}
    
  } else {
    message(taxa, ' has no data')}
}


``` 

\

# STEP 3 :: extract environmental values

\

Next we combine occurrence files from ALA and GBIF into one table, and extracts environmental values. It assumes that both files come from the combine_ala_records and combine_gbif_records functions. Note that the order of the raster names in 'world_raster' must match the order of names in the character vector 'env_variables'. In this case, it's simply the biolclim variables (i.e. bio1-bio19)

\

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=FALSE}

## The below shold really all be one big table
## Where we use different columns for the analysis :: species, genus or family
Insects_ALA      <- readRDS('./data/ALA/Insects/INVERTS_2022_1205.rds')


ALA.LAND.INV.SPP <- format_ala_dump(ALA_table     = Insects_ALA,
                                    record_type   = "ALA",
                                    keep_cols     = ALA_keep,
                                    year_filt     = FALSE,
                                    unique_cells  = FALSE,
                                    world_raster  = aus_annual_precip)

rm(Insects_ALA)



ALA_LAND_INV_SPP_sf <- SpatialPointsDataFrame(coords      = head(ALA.LAND.INV.SPP, 1000) %>% 
                                                dplyr::select(lon, lat) %>% as.matrix(),
                                              data        = head(ALA.LAND.INV.SPP, 1000),
                                              proj4string = CRS("+init=epsg:4326")) %>%
  
  st_as_sf() %>% 
  st_transform(., st_crs(4326))


ALA.LAND.INV.GENUS <- ALA.LAND.INV.SPP %>% dplyr::select(-searchTaxon) %>% 
  dplyr::rename(searchTaxon = genus)   %>% dplyr::select(searchTaxon, everything())


ALA.LAND.INV.FAM   <- ALA.LAND.INV.SPP %>% dplyr::select(-searchTaxon) %>% 
  dplyr::rename(searchTaxon = family)  %>% dplyr::select(searchTaxon, everything())



## Prepare site data 
## Combine all the site data into one table
all_insect_site_df_species <- PBI_AUS %>%
  
  ## Remove the NA coordinates, and add a 'SOURCE' column
  completeFun(., c('lat', 'lon')) %>% 
  filter(lon < 180 & lat > -90)   %>% 
  mutate(SOURCE = 'SITE')


## Create a SPDF for the SITE Data
all_insect_site_df_species_sf <- SpatialPointsDataFrame(coords      = all_insect_site_df_species %>% 
                                                          dplyr::select(lon, lat) %>% as.matrix(),
                                                        data        = all_insect_site_df_species,
                                                        proj4string = CRS("+init=epsg:4326")) %>%
  
  st_as_sf() %>% 
  st_transform(., st_crs(4326))


COMBO.RASTER.PBI.SPP = combine_records_extract(records_df       = all_insect_site_df_species_sf,
                                               add_sites        = FALSE,
                                               filter_taxo      = FALSE,
                                               
                                               site_df          = NA,
                                               thin_records     = FALSE,
                                               template_raster  = template_raster_250m,
                                               world_raster     = aus.climate.veg.grids.250m,
                                               epsg             = 4326,
                                               taxa_list        = target.insect.families,
                                               taxa_level       = 'family',   
                                               
                                               ## This might need to change too
                                               raster_divide    = FALSE,
                                               save_data        = FALSE,
                                               save_run         = FALSE)


COMBO_RASTER_PBI_SPP_sf <- SpatialPointsDataFrame(coords      = COMBO.RASTER.PBI %>% 
                                                    dplyr::select(lon, lat) %>% as.matrix(),
                                                  data        = COMBO.RASTER.PBI,
                                                  proj4string = CRS("+init=epsg:4326")) %>%
  
  st_as_sf() %>% 
  st_transform(., st_crs(4326))

st_write(COMBO_RASTER_PBI_SPP_sf, 
         dsn    = paste0(inv_records_dir, 'TEST_PBI_TAXA_POINTS.gpkg'), 
         quiet  = TRUE, 
         append = FALSE)



## Genera
COMBO.RASTER.PBI.GEN <- COMBO_RASTER_PBI_SPP_sf %>%
  dplyr::select(-searchTaxon) %>% 
  dplyr::rename(searchTaxon = genus) %>% dplyr::select(searchTaxon, everything())


## Familie
COMBO.RASTER.PBI.FAM <- COMBO_RASTER_PBI_SPP_sf  %>%
  dplyr::select(-searchTaxon)  %>% 
  dplyr::rename(searchTaxon = family) %>% dplyr::select(searchTaxon, everything())





## Update with the full insects download from ALA ----
## Combine GBIF and ALA data, and extract environmental values
## Some of the site records will be removed here, if they are outside the raster bounds.
## This seems strange, but maybe they are right on the coast
COMBO.RASTER.FAM = combine_records_extract(records_df       = ALA_LAND_INV_SPP_sf,
                                           add_sites        = FALSE,
                                           filter_taxo      = FALSE,
                                           
                                           site_df          = NA,
                                           thin_records     = FALSE,
                                           template_raster  = template_raster_250m,
                                           world_raster     = aus.climate.veg.grids.250m,
                                           epsg             = 4326,
                                           taxa_list        = target.insect.families,
                                           taxa_level       = 'family',   
                                           
                                           ## This might need to change too
                                           raster_divide    = FALSE,
                                           save_data        = FALSE,
                                           save_run         = "TARGET_INSECT_FAMILIES",
                                           data_path        = "./output/results/")
gc()


COMBO.RASTER.GEN = combine_records_extract(ala_df           = ALA.LAND.INV.GENUS,
                                           add_sites        = TRUE,
                                           filter_taxo      = FALSE,
                                           
                                           site_df          = all_insect_site_df_genera,
                                           thin_records     = FALSE,
                                           template_raster  = template_raster_250m,
                                           world_raster     = aus.climate.veg.grids.250m,
                                           prj              = CRS("+init=epsg:4326"),
                                           taxa_list        = target.insect.genera,
                                           taxa_level       = 'genus', 
                                           
                                           ## These two will need to change.
                                           ## Specify them in the code
                                           raster_divide    = FALSE,
                                           save_data        = FALSE,
                                           save_run         = "TARGET_INSECT_GENERA",
                                           data_path        = "./output/results/")
gc()


COMBO.RASTER.SPP = combine_records_extract(ala_df           = ALA.LAND.INV.SPP,
                                           add_sites        = TRUE,
                                           filter_taxo      = FALSE,
                                           
                                           site_df          = all_insect_site_df_species,
                                           thin_records     = FALSE,
                                           template_raster  = template_raster_250m,
                                           world_raster     = aus.climate.veg.grids.250m,
                                           prj              = CRS("+init=epsg:4326"),
                                           taxa_list        = target.insect.spp,
                                           taxa_level       = 'species', 
                                           
                                           ## These two will need to change.
                                           ## Specify them in the code
                                           raster_divide    = FALSE,
                                           save_data        = FALSE,
                                           save_run         = "TARGET_INSECT_SPECIES",
                                           data_path        = "./output/results/")
gc()


## Try combing the records into a "super table"...does that make sense?
COMBO.RASTER.FAM.GEN.SPP <- bind_rows(COMBO.RASTER.FAM, COMBO.RASTER.GEN, COMBO.RASTER.SPP)
gc()


COMBO_RASTER_ALA_SPP_sf <- SpatialPointsDataFrame(coords      = COMBO.RASTER.FAM %>% 
                                                    dplyr::select(lon, lat) %>% as.matrix(),
                                                  data        = COMBO.RASTER.FAM,
                                                  proj4string = CRS("+init=epsg:4326")) %>%
  
  st_as_sf() %>% 
  st_transform(., st_crs(4326))

st_write(COMBO_RASTER_ALA_SPP_sf, 
         dsn    = paste0(inv_records_dir, 'TEST_ALA_TAXA_POINTS.gpkg'), 
         quiet  = TRUE, 
         append = FALSE)


rm(COMBO.RASTER.FAM)
rm(COMBO.RASTER.GEN)
rm(COMBO.RASTER.SPP)
rm(ALA.LAND.INV.SPP)
rm(ALA.LAND.INV.GENUS)
rm(ALA.LAND.INV.FAM)




``` 

\

# STEP 4 :: Automated cleanin' of outlier records

\

The workflow uses four shapefiles as part of analysis and mapping: Australia, the World, 
the global Koppen Zones. The Koppen data are from CliMond, centred on 1975: 
https://www.climond.org/Core/Authenticated/KoppenGeiger.aspx

\

The next stage of the workflow use a series of cleaning functions to automate the removal 
of records for each taxa which are outliers. Doing this manually is extremely tedious, 
and although errors will be made, automation is preferable across large suites of taxa. 
The first cleaning function takes a data frame of all taxa records, and flag records 
as institutional or spatial outliers. This function uses the CoordinateCleaner package: 
https://cran.r-project.org/web/packages/CoordinateCleaner/index.html. It takes the 
records data.frame is that returned by the combine_records_extract function above.

\

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=FALSE}

## Flag records as institutional or spatial outliers
COORD.CLEAN = coord_clean_records(records    = COMBO.RASTER.FAM,
                                  site_flag  = 'SITE',
                                  occ_flag   = 'ALA',
                                  multi_source = TRUE,
                                  
                                  capitals   = 10000,  
                                  centroids  = 5000,   
                                  save_data  = FALSE,
                                  save_run   = "TARGET_INSECT_SPECIES",
                                  data_path  = "./output/results/")


SPATIAL.CLEAN = check_spatial_outliers(occ_df       = COORD.CLEAN,
                                       land_shp     = LAND,
                                       site_records = TRUE, 
                                       site_flag    = 'SITE',
                                       occ_flag     = 'ALA',
                                       multi_source = TRUE,
                                       clean_path   = './data/ALA/Check_Plots/Inverts/',
                                       plot_points  = FALSE,
                                       record_limit = 400000,
                                       spatial_mult = 10,
                                       prj          = CRS("+init=epsg:4326"))


GLOB.NICHE.ALL = calc_enviro_niches(coord_df     = SPATIAL.CLEAN %>% .[.$searchTaxon %in% target.insect.families, ],
                                    prj          = CRS("+init=epsg:4326"),
                                    country_shp  = AUS,
                                    world_shp    = LAND,
                                    kop_shp      = Koppen_shp,
                                    taxa_list    = analysis_taxa,
                                    env_vars     = names(aus.climate.veg.grids.250m),
                                    cell_size    = 2,
                                    save_data    = TRUE,
                                    save_run     = "ALL_INSECT_FAMILIES_ALA_PBI",
                                    data_path    = inv_results_dir)


plot_range_histograms(coord_df     = SPATIAL.CLEAN %>% .[.$searchTaxon %in% target.insect.families, ],
                      taxa_list    = target.insect.genera,
                      range_path   = check_dir)

``` 



# STEP 5 :: Prepare SDM table

\

Then we need to create at table we can use for taxa distribution modelling. This function takes 
a data frame of all taxa records, and prepares a table in the 'taxa with data' (swd) format for 
modelling uses the Maxent algorithm. It assumes that the input df is that returned by the 
coord_clean_records function.

\

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=FALSE}

## Only use GDA Albers forAus analyses
sp_epsg3577  <- "+proj=aea +lat_0=0 +lon_0=132 +lat_1=-18 +lat_2=-36 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"


COORD.CLEAN = coord_clean_records(records    = COMBO.RASTER.FAM,
                                  site_flag  = 'SITE',
                                  occ_flag   = 'ALA',
                                  multi_source = TRUE,
                                  
                                  capitals   = 10000,  
                                  centroids  = 5000,   
                                  save_data  = FALSE,
                                  save_run   = "TARGET_INSECT_SPECIES",
                                  data_path  = "./output/results/")


## Combine occ data with the bg data 
SDM.SPAT.OCC.BG.GDA <- prepare_sdm_table(coord_df          = COORD.CLEAN,
                                         taxa_list         = analysis_taxa,
                                         
                                         site_flag         = 'SITE',
                                         occ_flag          = 'ALA',
                                         site_records      = TRUE,
                                         
                                         sdm_table_vars    = c('searchTaxon', 
                                                               'species',  
                                                               'genus', 
                                                               'family',
                                                               'order',
                                                               'class',                         
                                                               'phylum',                        
                                                               'kingdom',
                                                               'lon',      
                                                               'lat',
                                                               'year',
                                                               'SOURCE', 
                                                               'SPAT_OUT',
                                                               names(aus.climate.veg.grids.250m)),
                                         
                                         save_run          = "TEST_INVERTS_ALA_PBI",
                                         read_background   = FALSE,
                                         sp_country_prj    = sp_epsg3577,
                                         save_data         = TRUE,
                                         data_path         = inv_results_dir)


## Check the taxonomic levels
sum(is.na(SDM.SPAT.OCC.BG.GDA@data$searchTaxon));sum(is.na(SDM.SPAT.OCC.BG.GDA@data$species));
sum(is.na(SDM.SPAT.OCC.BG.GDA@data$genus));sum(is.na(SDM.SPAT.OCC.BG.GDA@data$family))


## Create subset of target reptiles
## Save each taxa as an individual shapefile
# SDM.SPAT.OCC.BG.GDA = readRDS('./output/invert_maxent_raster_update/results/SDM_SPAT_OCC_BG_ALL_TARGET_INSECT_TAXA.rds')
SDM.SPAT.OCC.BG.TARG.INV <- SDM.SPAT.OCC.BG.GDA %>% .[.$searchTaxon %in% analysis_taxa, ]


st_write(SDM.SPAT.OCC.BG.TARG.INV %>% st_as_sf(), 
         dsn = file.path(inv_results_dir, 'SDM_INVERT_TEST_ALA_PBI.gpkg'), 
         layer = 'SDM_TARGET_INVERT_TAXA', 
         quiet = TRUE)


# st_write(SDM.SPAT.OCC.BG.GDA %>% st_as_sf(), 
#          dsn = file.path(getwd(), 'output/invert_maxent_raster_update/results/SDM_INVERT_ALL_TAXA_ALA_PBI.gpkg'), 
#          layer = 'SDM_ALL_INVERT_TAXA', 
#          quiet = TRUE)

``` 

\

# Check SDM data

\

We need to check that all the taxa analysed have all the quality records we need.

\

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=FALSE}

## Save each taxa as an individual shapefile
SDM.SPAT.OCC.BG.GDA      <- readRDS(paste0(inv_results_dir, 'SDM_SPAT_OCC_BG_ALL_TARGET_INSECT_TAXA_UPDATE.rds'))


## Also save a big table of just the background taxa
SDM.SPAT.OCC.BG.TARG.GDA <- SDM.SPAT.OCC.BG.GDA %>% .[.$searchTaxon %in% analysis_taxa, ]
SDM.SPAT.OCC.BG.TARG.FAM <- SDM.SPAT.OCC.BG.GDA %>% .[.$searchTaxon %in% target.insect.families, ]
SDM.SPAT.OCC.BG.TARG.GEN <- SDM.SPAT.OCC.BG.GDA %>% .[.$searchTaxon %in% target.insect.genera, ]
SDM.SPAT.OCC.BG.TARG.SPP <- SDM.SPAT.OCC.BG.GDA %>% .[.$searchTaxon %in% target.insect.spp, ]


st_write(SDM.SPAT.OCC.BG.TARG.FAM %>% st_as_sf(), 
         dsn   = file.path(getwd(), 
                           paste0(inv_results_dir, 'SDM_ALL_INVERT_TAXA_ALA_PBI.gpkg')), 
         layer = 'SDM_TARGET_INVERT_FAMILIES', 
         quiet = TRUE)


st_write(SDM.SPAT.OCC.BG.TARG.GEN %>% st_as_sf(), 
         dsn   = file.path(getwd(), paste0(inv_results_dir, 'SDM_ALL_INVERT_GENERA_ALA_PBI.gpkg')), 
         layer = 'SDM_TARGET_INVERT_GENERA', 
         quiet = TRUE)


st_write(SDM.SPAT.OCC.BG.TARG.SPP %>% st_as_sf(), 
         dsn   = file.path(getwd(), paste0(inv_results_dir, 'SDM_ALL_INVERT_SPECIES_ALA_PBI.gpkg')), 
         layer = 'SDM_TARGET_INVERT_SPECIES', 
         quiet = TRUE)


## Loop through families
for(taxa in target.insect.families) {
  
  ## 
  if(taxa %in% unique(SDM.SPAT.OCC.BG.TARG.GDA$searchTaxon)) {
    
    taxa_shp <- paste0(inv_records_dir,
                       taxa, '_SDM_ALA_PBI_points.shp')
    
    if(!file.exists(taxa_shp)) {
      
      message('Subsetting ', taxa, ' shapefile and geo-package layers')
      taxa_occ <- SDM.SPAT.OCC.BG.TARG.GDA %>% .[.$searchTaxon %in% taxa, ]
      
      st_write(taxa_occ %>% st_as_sf(), 
               paste0(inv_records_dir, taxa, '_SDM_ALA_PBI_POINTS.shp'))
      
      st_write(taxa_occ %>% st_as_sf(), 
               dsn = paste0(inv_records_dir, 'SDM_INVERT_TARG_TAXA_ALA_PBI.gpkg'), 
               layer = paste0(taxa, '_SDM_points'), 
               quiet = TRUE)
      
    } else {
      message(taxa, ' SDM .shp already exists')}
    
  } else {
    message(taxa, ' has no data')}
}



## Loop through genera
for(taxa in target.insect.genera) {
  
  ## 
  if(taxa %in% unique(SDM.SPAT.OCC.BG.TARG.GDA$searchTaxon)) {
    
    taxa_shp <- paste0(inv_records_dir,
                       taxa, '_SDM_ALA_PBI_points.shp')
    
    if(!file.exists(taxa_shp)) {
      
      message('Subsetting ', taxa, ' shapefile and geo-package layers')
      taxa_occ <- SDM.SPAT.OCC.BG.TARG.GDA %>% .[.$searchTaxon %in% taxa, ]
      
      st_write(taxa_occ %>% st_as_sf(), 
               paste0(inv_records_dir, taxa, '_SDM_ALA_PBI_points.shp'))
      
      st_write(taxa_occ %>% st_as_sf(), 
               dsn = paste0(inv_records_dir, 'SDM_INVERT_TARG_TAXA__ALA_PBI.gpkg'), 
               layer = paste0(taxa, '_SDM_points'), 
               quiet = TRUE)
      
    } else {
      message(taxa, ' SDM .shp already exists')}
    
  } else {
    message(taxa, ' has no data')}
}





## Loop through species
for(taxa in target.insect.spp) {
  
  ## taxa = target.insect.spp[79] 
  if(taxa %in% unique(SDM.SPAT.OCC.BG.TARG.GDA$searchTaxon)) {
    
    taxa_shp <- paste0(inv_records_dir,
                       taxa, '_SDM_ALA_PBI_points.shp')
    
    if(!file.exists(taxa_shp)) {
      
      message('Subsetting ', taxa, ' shapefile and geo-package layers')
      taxa_occ <- SDM.SPAT.OCC.BG.TARG.GDA %>% .[.$searchTaxon %in% taxa, ]
      
      st_write(taxa_occ %>% st_as_sf(), 
               paste0(inv_records_dir, taxa, '_SDM_ALA_PBI_points.shp'))
      
      st_write(taxa_occ %>% st_as_sf(), 
               dsn = paste0(inv_records_dir, taxa, '_SDM_ALA_PBI_points.gpkg'), 
               layer = paste0(taxa, '_SDM_points'), 
               quiet = TRUE)
      
      st_write(taxa_occ %>% st_as_sf(), 
               dsn = paste0(inv_records_dir, 'SDM_INVERT_TARG_TAXA_ALA_PBI.gpkg'), 
               layer = paste0(taxa, '_SDM_points'), 
               quiet = TRUE)
      
    } else {
      message(taxa, ' SDM .shp already exists')}
    
  } else {
    message(taxa, ' has no data')}
}


``` 

\
