---
title: "nenswniche : a package for rapidly estimating multiple taxa ranges and habitat suitability"
authors: "Hugh Burley, Shawn Laffan, Will Cornwell, Adrian Fisher"
date: "October 2021"
output:
  github_document:
  toc: true   
toc_depth: 4            
toc_float: true
number_sections: false  
vignette: >
%\VignetteIndexEntry{README}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
  keep_md: true
  theme: united     
highlight: tango        
css: styles.css
revealjs::revealjs_presentation:
  dev: 'svg'
chunk_output_type: console
self_contained: false
reveal_plugins: ["notes", "search"]
reveal_options:
  slideNumber: true
previewLinks: true
word_document:
  always_allow_html: yes
---

\

The text and code below summarises a workflow in R that can be used to relatively rapidly assess the environmental range of a taxa within Australia, from downloading occurrence records, through to creating maps of predicted climatic suitability across Australia at 1km*1km resolution. An example of this work is published in the journal Science of the Total Environment ::

\

Burley, H., Beaumont, L.J., Ossola, A., et al. (2019) Substantial declines in urban tree habitat predicted 
under climate change. Science of The Total Environment, 685, 451-462.

https://www.sciencedirect.com/science/article/pii/S0048969719323289#f0030 

\

To install, run :

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=FALSE}

## Set environments 
## java memory limit and temporary raster dir
rm(list = ls())
options(java.parameters = "-Xmx64000m")


## Function to load or install packages
ipak <- function(pkg){
  
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg))
    install.packages(new.pkg, dependencies = TRUE, 
                     repos = "https://cran.csiro.au/")
  
  sapply(pkg, require, character.only = TRUE)
  
}

library(nenswniche)
data('sdmgen_packages')
ipak(sdmgen_packages)


## Set temporary raster dir for the terra package
rasterOptions(tmpdir = 'G:/GITHUB/North_east_NSW_fire_recovery/TEMP')
terraOptions(memfrac = 0.5, 
             tempdir = 'G:/GITHUB/North_east_NSW_fire_recovery/TEMP')

```

\
\
\

# Background

This code is being developed at UNSW, as part of a project investigating
the impacts of the 2019/2020 bush fires on Insects in the North East Forests 
of New South Wales. The aim is to create a pipeline that rapidly assesses 
the habitat suitability of the threatened insect taxa under current 
environmental conditions. 

\

# STEP 1 :: Download taxa occurrence data

\

The backbone of the R workflow is a list of (taxonomically Ridgey-Didge!) Taxa names 
that we supply. The analysis is designed to process data for one taxa at a time, 
allowing taxa results to be updated as required. Unfortunately, Australia's insects
are not very well sampled...so we can analyse at the family level.

\

Let's use all the insect families that we think might be threatened by the 2019/2020 fires. 

\

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=TRUE}


## To Do :
## 1). Check the errors that Finlay may have found
## 2). Clean up the folders


## Get the invertebrate site data
source('./R/CREATE_SPECIES_LIST.R')
source('./R/SDM_GEN_PROCESSOR_FUNCTIONS.R')
source('./R/SDM_GEN_MAXENT_FUNCTIONS.R')
source('./R/SDM_GEN_MAPPING_FUNCTIONS.R')


## Create lists of all taxa that we will model at all levels
data('target.host.plants')
data('all.insect.plant.spp')


```


The taxa list is supplied to a series of functions to calculate environmental ranges and habitat 
suitability. The initial functions download all taxa records from the Atlas and living Australia
(https://www.ala.org.au/) and the Global Biodiversity Information Facility (GBIF, https://www.gbif.org/). 
The taxa data are downloaded as individual .Rdata files to the specified folders, which must exist first, 
without returning anything.

\

Skip this step for data test
Now download GBIF and ALA occurrence data for each taxa. The downloading functions 
are separated, because the ALA and GBIF columns are slightly different, but both 
data sources are needed to properly quantify taxa ranges. The package functions 
expect these folders (a typical R project structure), create them if they don't exist

\

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=FALSE}

## Download ALA occurrence data for each taxa
## For some reason, this temporary directory needed to be created...
## dir.create('C:/Users/hughb/AppData/Local/Temp/RtmpyijnUR')

## Get some extra columns from the ALA with environmental data
## This could be better organized by getting the column names by 'category' - enviro, land context, etc
ala_env_cols = ala_fields("occurrence_stored", as_is = TRUE)$description[231:389]


##
download_ALA_all_species(species_list   = all.insect.plant.spp,
                         your_email     = 'hugh.burley@gmail.com',
                         download_path  = "./data/ALA/Plants/",
                         ala_temp_dir   = 'C:/Users/hughb/AppData/Local/Temp/RtmpCmco7x',
                         download_limit = 20000,
                         # extra_cols     = ala_env_cols,
                         quality_cols   = 'all')


``` 

\

# STEP 2 :: Combine taxa occurrence data

\

This pipeline combines taxa occurrence points with raster data. It was developed using worldclim 
climate raster data, but it can take any set of rasters. All the raster data is available here :: 

\

https://drive.google.com/open?id=1T5ET5MUX3-lkqiN5nNL3SZZagoJlEOal. 

\

Let's create all the raster data we need. First let's get the rasters needed for SDM modelling :
- Climate
- Vegetation
- Topography
- Soil


\

# 250m Climate, Soil and terrain variables

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=FALSE}

## 250m Precip layers
aus_precip_250m <- raster::stack(
  list.files('./data/Bushfire_indices/R_outputs/250m/AUS/Precip',        pattern =".tif", full.names = TRUE))

east_precip_250m <- raster::stack(
  list.files('./data/Bushfire_indices/R_outputs/250m/EAST_COAST/Precip', pattern =".tif", full.names = TRUE))


## 250m temperature layers
aus_temp_250m <- raster::stack(
  list.files('./data/Bushfire_indices/R_outputs/250m/AUS/Temp',          pattern =".tif", full.names = TRUE))

east_temp_250m <- raster::stack(
  list.files('./data/Bushfire_indices/R_outputs/250m/EAST_COAST/Temp',   pattern =".tif", full.names = TRUE))


## 250m Soil layers
aus_soil_250m <- raster::stack(
  list.files('./data/Bushfire_indices/R_outputs/250m/AUS/Soil',          pattern =".tif", full.names = TRUE))

east_soil_250m <- raster::stack(
  list.files('./data/Bushfire_indices/R_outputs/250m/EAST_COAST/Soil',   pattern =".tif", full.names = TRUE))


## 250m Geology Australia
aus_geology_250m <- raster::stack(
  list.files('./data/Bushfire_indices/R_outputs/250m/AUS/Geology',        pattern =".tif", full.names = TRUE))

east_geology_250m <- raster::stack(
  list.files('./data/Bushfire_indices/R_outputs/250m/EAST_COAST/Geology', pattern =".tif", full.names = TRUE))


## 250m topo Australia
aus_topo_250m <- raster::stack(
  list.files('./data/Bushfire_indices/R_outputs/250m/AUS/Topo',           pattern =".tif", full.names = TRUE))

east_topo_250m <- raster::stack(
  list.files('./data/Bushfire_indices/R_outputs/250m/EAST_COAST/Topo',    pattern =".tif", full.names = TRUE))


## 250m terrain indices Australia
aus_terrain_250m <- raster::stack(
  list.files('./data/Bushfire_indices/R_outputs/250m/AUS/Indices',        pattern =".tif", full.names = TRUE))

east_terrain_250m <- raster::stack(
  list.files('./data/Bushfire_indices/R_outputs/250m/EAST_COAST/Indices', pattern =".tif", full.names = TRUE))


## Climate data for Australia, in GDA Albers projection
aus.grids.current.250m <- stack(aus_precip_250m,
                                aus_temp_250m,
                                aus_soil_250m,
                                aus_topo_250m,
                                aus_terrain_250m,
                                aus_geology_250m)

east.grids.current.250m <- stack(east_precip_250m,
                                 east_temp_250m,
                                 east_soil_250m,
                                 east_topo_250m,
                                 east_terrain_250m,
                                 east_geology_250m)


## And a stack of grids for vegetation, in GDA Albers projection
aus.veg.grids.250m <- stack(
  
  list.files('./data/Remote_sensing/Veg_data/height_and_cover/Aus/250m', 
             '_250m.tif', full.names = TRUE))

east.veg.grids.250m <- stack(
  
  list.files('./data/Remote_sensing/Veg_data/height_and_cover/Eastern_Aus/250m', 
             '_east_coast.tif', full.names = TRUE))


names(east.grids.current.250m) <- gsub('_EAST_COAST', '', names(east.grids.current.250m))
names(aus.veg.grids.250m)      <- names(east.veg.grids.250m) <- c("Plant_cover_fraction_0_5m", 
                                                                  "Plant_cover_fraction_5_10m",  
                                                                  "Plant_cover_fraction_10_30m",      
                                                                  "Plant_cover_fraction_30m",
                                                                  "Total_Plant_cover_fraction",  
                                                                  "Tree_canopy_height_25th", 
                                                                  "Tree_canopy_height_50th", 
                                                                  "Tree_canopy_height_75th",   
                                                                  "Tree_canopy_height_95th",   
                                                                  "Tree_canopy_peak_foliage",
                                                                  "Tree_canopy_peak_foliage_total",
                                                                  "mrvbf")


## Combine the grids into raster stacks
aus.climate.veg.grids.250m   <- stack(aus.grids.current.250m, aus.veg.grids.250m)
east.climate.veg.grids.250m  <- stack(east.grids.current.250m, east.veg.grids.250m)
aus_annual_precip            <- raster('./data/Bushfire_indices/R_outputs/250m/AUS/Extra/Annual_precip_WGS84.tif')
aus_annual_precip_alb        <- raster('./data/Bushfire_indices/R_outputs/250m/AUS/Extra/Annual_precip_GDA_ALB.tif')


## Should be 1km*1km, It should have a value of 1 for land, and NA for the ocean
aus_annual_precip_alb[aus_annual_precip_alb > 0] <- 1
template_raster_250m <- aus_annual_precip_alb


## Add shapefiles
NENSW_study_area <- readOGR(dsn              = './data/Spatial_data/NE_NSW_study_area_buffer_GDA.shp',
                            layer            = "NE_NSW_study_area_buffer_GDA",
                            stringsAsFactors = FALSE)


east_study_area <- readOGR(dsn              = './data/Spatial_data/east_coast_shapefile/polygon_east_coast.shp',
                           layer            = "polygon_east_coast",
                           stringsAsFactors = FALSE)


```


\

Next we filter the records to those taken after 1950, and those inside the raster boundaries 
(i.e. taxa records in the ocean according to the raster boundaries will be excluded).

\

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=FALSE}

## Combine ALA data, and filter to records on land taken > 1950
## The climate data is the worldclim version 2.0
## Also, add a switch for the 1km filter - we don't want to do this anymore, too few records
data('ALA_keep')


## Combine  
ALA.LAND.PLA <- combine_ala_records(taxa_list         = all.insect.plant.spp,
                                    records_path      = "./data/ALA/Plants/",
                                    records_extension = "_ALA_records.RData",
                                    record_type       = "ALA",
                                    keep_cols         = ALA_keep,
                                    world_raster      = aus_annual_precip)

``` 

\

# STEP 3 :: extract environmental values

\

Next we combine occurrence files from ALA and GBIF into one table, and extracts environmental values. It assumes that both files come from the combine_ala_records and combine_gbif_records functions. Note that the order of the raster names in 'world_raster' must match the order of names in the character vector 'env_variables'. In this case, it's simply the biolclim variables (i.e. bio1-bio19)

\

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=FALSE}

## Combine GBIF and ALA data, and extract environmental values
## Some of the site records will be removed here, if they are outside the raster bounds.
## This seems strange, but maybe they are right on the coast
COMBO.RASTER.SPP <- combine_records_extract(ala_df          = ALA.LAND.PLA,
                                           add_sites        = FALSE,
                                           filter_taxo      = FALSE,
                                           
                                           site_df          = FALSE,
                                           thin_records     = FALSE,
                                           template_raster  = template_raster_250m,
                                           world_raster     = aus.climate.veg.grids.250m,
                                           prj              = CRS('+init=epsg:4326'),
                                           taxa_list        = all.insect.plant.spp[1:3],
                                           taxa_level       = 'species',   
                                           
                                           ## This might need to change too
                                           raster_divide    = FALSE,
                                           save_data        = FALSE,
                                           save_run         = 'ALL_PLANT_TAXA',
                                           data_path        = './output/plant_maxent_raster_update/results/')


``` 

\

# STEP 4 :: Automated cleanin' of outlier records

\

The workflow uses four shapefiles as part of analysis and mapping: Australia, the World, 
the global Koppen Zones. The Koppen data are from CliMond, centred on 1975: 
https://www.climond.org/Core/Authenticated/KoppenGeiger.aspx

\

The next stage of the workflow use a series of cleaning functions to automate the removal 
of records for each taxa which are outliers. Doing this manually is extremely tedious, 
and although errors will be made, automation is preferable across large suites of taxa. 
The first cleaning function takes a data frame of all taxa records, and flag records 
as institutional or spatial outliers. This function uses the CoordinateCleaner package: 
https://cran.r-project.org/web/packages/CoordinateCleaner/index.html. It takes the 
records data.frame is that returned by the combine_records_extract function above.

\

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=FALSE}

## Just create a background points version for each step 

## Flag records as institutional or spatial outliers
COORD.CLEAN = coord_clean_records(records    = COMBO.RASTER.SPP,
                                  site_flag  = 'SITE',
                                  occ_flag   = 'ALA',
                                  multi_source = TRUE,
                                  
                                  capitals   = 10000,  
                                  centroids  = 5000,   
                                  save_data  = FALSE,
                                  save_run   = "TARGET_INSECT_SPECIES",
                                  data_path  = "./output/results/")

``` 

\

The next cleaning function takes a data frame of all taxa records, flags records as 
spatial outliers (T/F for each record in the df), and saves images of the checks for each. 
Manual cleaning of spatial outliers is very tedious, but automated cleaning makes mistakes, 
so checking is handy. This function uses the CoordinateCleaner package 
https://cran.r-project.org/web/packages/CoordinateCleaner/index.html. It assumes that the 
input dfs are those returned by the coord_clean_records function.

\

The next cleaning function takes a data frame of all taxa records, estimates the geographic 
and environmental ranges for each taxa, and creates a table of all taxa ranges. 
It uses the AOO.computing function in the ConR package: https://cran.r-project.org/web/packages/ConR/index.html
It assumes that the input df is that returned by the check_spatial_outliers function.

\


# STEP 5 :: Prepare SDM table

\

Then we need to create at table we can use for taxa distribution modelling. This function takes 
a data frame of all taxa records, and prepares a table in the 'taxa with data' (swd) format for 
modelling uses the Maxent algorithm. It assumes that the input df is that returned by the 
coord_clean_records function.

\

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=FALSE}

## Only use GDA Albers
sp_epsg3577  <- "+proj=aea +lat_0=0 +lon_0=132 +lat_1=-18 +lat_2=-36 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"


## Combine occ data with the bg data 
SDM.SPAT.OCC.BG <- prepare_sdm_table(coord_df          = COORD.CLEAN,
                                     taxa_list         = analysis_taxa,
                                     
                                     site_flag         = 'SITE',
                                     occ_flag          = 'ALA',
                                     site_records      = FALSE,
                                     
                                     sdm_table_vars    = c('searchTaxon', 
                                                           'species',  
                                                           'genus', 
                                                           'family',      
                                                           'lon',      
                                                           'lat',
                                                           'year',
                                                           'SOURCE', 
                                                           'SPAT_OUT',
                                                           names(aus.climate.veg.grids.250m)),
                                     
                                     save_run          = "ALL_TARGET_HOST_PLANTS",
                                     read_background   = FALSE,
                                     sp_country_prj    = sp_epsg3577,
                                     save_data         = TRUE,
                                     data_path         = paste0(getwd(), '/output/plant_maxent_raster_update/results/'),
                                     project_path      = 'G:/GITHUB/North_east_NSW_fire_recovery')


## Check the taxonomic levels
sum(is.na(SDM.SPAT.OCC.BG@data$searchTaxon));sum(is.na(SDM.SPAT.OCC.BG@data$species));
sum(is.na(SDM.SPAT.OCC.BG@data$genus));sum(is.na(SDM.SPAT.OCC.BG@data$family))


## Create subset of target reptiles
SDM.SPAT.OCC.BG.TARG.INV <- SDM.SPAT.OCC.BG %>% .[.$searchTaxon %in% target.host.plants, ]


## Save shape file separately - doesn't seem to work in R
## Save the data to a shapefile
writeOGR(obj    = SDM.SPAT.OCC.BG,
         dsn    = 'G:/Github_repos/North_east_NSW_fire_recovery/output/plant_maxent_raster_update/results',
         layer  = 'SDM_TARGET_PLANT_TAXA',
         driver = 'ESRI Shapefile', overwrite_layer = TRUE)


st_write(SDM.SPAT.OCC.BG.TARG.INV %>% st_as_sf(), 
         dsn = file.path(getwd(), 'output/plant_maxent_raster_update/results/SDM_PLANT_TAXA.gpkg'), 
         layer = 'SDM_ALL_PLANT_TAXA', 
         quiet = TRUE)


st_write(SDM.SPAT.OCC.BG %>% st_as_sf(), 
         dsn = file.path(getwd(), 'output/plant_maxent_raster_update/results/SDM_PLANT_TAXA.gpkg'), 
         layer = 'SDM_TARGET_PLANT_TAXA', 
         quiet = TRUE)

``` 

\
