---
title: "nenswniche : a package for rapidly estimating multiple species ranges and habitat suitability"
authors: "Hugh Burley, Shawn Laffan, Adrian Fisher"
date: "March 2021"
output:
  github_document:
  toc: true   
toc_depth: 4            
toc_float: true
number_sections: false  
vignette: >
%\VignetteIndexEntry{README}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
  keep_md: true
  theme: united     
highlight: tango        
css: styles.css
revealjs::revealjs_presentation:
  dev: 'svg'
chunk_output_type: console
self_contained: false
reveal_plugins: ["notes", "search"]
reveal_options:
  slideNumber: true
previewLinks: true
word_document:
  always_allow_html: yes
---

\

The text and code below summarises a workflow in R that can be used to relatively rapidly assess the environmental range of a species within Australia, from downloading occurrence records, through to creating maps of predicted climatic suitability across Australia at 1km*1km resolution. An example of this work is published in the journal Science of the Total Environment ::

\

Burley, H., Beaumont, L.J., Ossola, A., et al. (2019) Substantial declines in urban tree habitat predicted 
under climate change. Science of The Total Environment, 685, 451-462.

https://www.sciencedirect.com/science/article/pii/S0048969719323289#f0030 

\

To install, run :

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=FALSE}

## Function to load or install packages
ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg))
    install.packages(new.pkg, dependencies = TRUE, repos="https://cran.csiro.au/")
  sapply(pkg, require, character.only = TRUE)
}

## These packages are not on cran, but are needed
# library(devtools)
# install_github('johnbaums/things')
# install_github('johnbaums/rmaxent')
# install_github('traitecoevo/taxonlookup')

## Main package 
# devtools::install_github("HMB3/nenswniche")

## Load packages 
library(nenswniche)
data('sdmgen_packages')
ipak(sdmgen_packages)

```

\
\
\

# Background

This code is being developed at UNSW, as part of a project investigating
the impacts of the 2019/2020 bush fires on Insects in the North East Forests 
of New South Wales. The aim is to create a pipeline that rapidly assesses 
the habitat suitability of the threatened insect species under current 
environmental conditions. 

\

# STEP 1 :: Download species occurrence data

\

The backbone of the R workflow is a list of (taxonomically Ridgey-Didge!) Taxa names 
that we supply. The analysis is designed to process data for one species at a time, 
allowing species results to be updated as required. Unfortunately, Australia's insects
are not very well sampled...so we can analyse at the family level.

\

Let's use all the insect families that we think might be threatened by the 2019/2020 fires. 

\

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=TRUE}

## Create lists of all taxa that we will model at all levels
data('insect_data_families')
data('all.insect.families')
data('all.insect.genera')
data('all.insect.spp')

data('target.insect.spp')
data('target.insect.genera')
data('target.insect.families')
data('target.host.plants')
data('all.insect.plant.spp')

analysis_taxa <- str_trim(c(target.insect.spp, target.insect.genera, target.insect.families)) %>% unique() 
#analysis_taxa <- str_trim(c(target.insect.spp, all.insect.spp)) %>% unique() 
analysis_taxa

```


The species list is supplied to a series of functions to calculate environmental ranges and habitat 
suitability. The initial functions download all species records from the Atlas and living Australia
(https://www.ala.org.au/) and the Global Biodiversity Information Facility (GBIF, https://www.gbif.org/). 
The species data are downloaded as individual .Rdata files to the specified folders, which must exist first, 
without returning anything.

\

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=FALSE}

## The functions expect these folders,
ALA_dir     <- './data/ALA'
GBIF_dir    <- './data/GBIF'

back_dir     <- './output/climate_maxent/back_sel_models'
full_dir     <- './output/climate_maxent/full_models'

comb_back_dir <- './output/full_maxent/back_sel_models'
comb_full_dir <- './output/full_maxent/full_models'

veg_back_dir <- './output/veg_maxent/back_sel_models'
veg_full_dir <- './output/veg_maxent/full_models'

results_dir <- './output/results'
climate_dir <- './data/worldclim/world/2070'
check_dir   <- './data/GBIF/Check_plots/'
temp_dir    <- 'G:/Raster_temp'

dir_lists   <- c(ALA_dir,  GBIF_dir,    back_dir, check_dir, veg_back_dir, veg_full_dir,
                 full_dir, results_dir, climate_dir, temp_dir)

## Create the folders if they don't exist
for(i in dir_lists) {
  if(!dir.exists(i)) {
    message('Creating ', i, ' directory')
    dir.create(i) } else {
      message(i, ' directory already exists')}
}

```

\

Now download GBIF and ALA occurrence data for each species. The downloading functions 
are separated, because the ALA and GBIF columns are slightly different, but both 
data sources are needed to properly quantify species ranges. The package functions 
expect these folders (a typical R project structure), create them if they don't exist

\

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=FALSE}

## Download ALA occurrence data for each species
## For some reason, this temporary directory needed to be created...
## dir.create('C:/Users/hughb/AppData/Local/Temp/RtmpyijnUR')

## Get some extra columns from the ALA with environmental data
## This could be better organized by getting the column names by 'category' - enviro, land context, etc
ala_env_cols = ala_fields("occurrence_stored", as_is = TRUE)$description[231:389]


##
download_ALA_all_species(species_list   = analysis_taxa,
                         your_email     = 'hugh.burley@gmail.com',
                         download_path  = "./data/ALA/Insects/Species/",
                         ala_temp_dir   = 'C:/Users/hughb/AppData/Local/Temp/RtmpCmco7x',
                         download_limit = 20000,
                         extra_cols     = ala_env_cols,
                         quality_cols   = 'all')


download_ALA_all_genera(species_list    = analysis_taxa,
                        your_email      = 'hugh.burley@gmail.com',
                        download_path   = "./data/ALA/Insects/Genus/",
                        ala_temp_dir    = 'C:/Users/hughb/AppData/Local/Temp/RtmpCmco7x',
                        download_limit  = 20000,
                        extra_cols      = ala_env_cols,
                        quality_cols    = 'all')


## Now download the data for all the families
download_ALA_all_families(species_list   = analysis_taxa,
                          your_email     = 'hugh.burley@gmail.com',
                          download_path  = "./data/ALA/Insects/Family/",
                          ala_temp_dir   = 'C:/Users/hughb/AppData/Local/Temp/RtmpCmco7x',
                          download_limit = 20000,
                          extra_cols     = ala_env_cols,
                          quality_cols   = 'all')

``` 

\

# STEP 2 :: Combine species occurrence data

\

This pipeline combines species occurrence points with raster data. It was developed using worldclim 
climate raster data, but it can take any set of rasters. All the raster data is available here :: 

\

https://drive.google.com/open?id=1T5ET5MUX3-lkqiN5nNL3SZZagoJlEOal. 

\

Let's create all the raster data we need here :

\

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=FALSE}

## Use worldclim 2 data
## https://drive.google.com/open?id=1mQHVmYxSMw_cw1iGvfU9M7Pq6Kl6nz-C
worldclim_climate = raster::stack(
  file.path('./data/worldclim2/res_30s/bioclim',
            sprintf('wc2.1_30s_bio_%d.tif', 1:19)))

## Climate data for Australia, in GDA Albers projection
aus.grids.current <- stack(
  file.path('./data/worldclim2/aus/current', 
            sprintf('bio_%d_AUSGDA.tif', 1:19)))

## Climate data for the study area in NE_NSW, in GDA Albers projection
study.grids.current <- stack(
  file.path('./data/worldclim2/aus/current/Study_area',
            sprintf('bio_%d_AUSGDA.tif', 1:19)))

## And a stack of grids for vegetation, in GDA Albers projection
aus.veg.grids <- stack(
  list.files('./data/Remote_sensing/Veg_data/height_and_cover', 
             '_1km.tif', full.names = TRUE))

## Rename the raster grids
names(worldclim_climate)   <- env_variables
names(aus.grids.current)   <- names(worldclim_climate)
names(study.grids.current) <- names(worldclim_climate)


names(aus.veg.grids) <- c("Plant_cover_fraction_0_5m", 
                          "Plant_cover_fraction_5_10m",  
                          "Plant_cover_fraction_10_30m",      
                          "Plant_cover_fraction_30m",
                          "Total_Plant_cover_fraction",  
                          "Tree_canopy_height_25th", 
                          "Tree_canopy_height_50th", 
                          "Tree_canopy_height_75th",   
                          "Tree_canopy_height_95th",   
                          "Tree_canopy_peak_foliage",
                          "Tree_canopy_peak_foliage_total")

## Combine the grids
aus.climate.veg.grids <- stack(aus.grids.current, aus.veg.grids)
aus_annual_temp       <- aus.climate.veg.grids[[1]]
world_annual_temp     <- worldclim_climate[[1]]


# We also need a template raster of 1km * 1km cells, which is used to filter 
# records to 1 per one 1km cell. This raster needs to have the same extent (global) 
# resolution (1km) and projection (WGS84) of the data used to analyse the species 
# distributions. It should have a value of 1 for land, and NA for the ocean. 
# This takes ages in R.....
template_raster_1km_WGS84 = raster("./data/world_koppen/template_1km_WGS84.tif")
Koppen_1975_1km           = raster('data/world_koppen/Koppen_1000m_Mollweide54009.tif')
template_raster_1km_mol   = raster("./data/world_koppen/template_has_data_1km.tif")


## Use Molleweide or GDA Albers
# sp_epsg3577 <- "+proj=aea +lat_0=0 +lon_0=132 +lat_1=-18 +lat_2=-36 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs"
# sp_epsg54009 <- '+proj=moll +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs +towgs84=0,0,0'

## Use gdal to create a template raster in the mollweide projection, using one of the bioclim layers
# template_raster_1km_GDA <- gdalwarp("./data/worldclim2/aus/current/bio_1_AUSGDA.tif",
#                                     tempfile(fileext = '.tif'),
#                                     t_srs = sp_epsg3577,
#                                     output_Raster = TRUE,
#                                     tr = c(1000, 1000),
#                                     r = "near", 
#                                     dstnodata = '-9999')

## Should be 1km*1km, values of 0 for ocean and 1 for land 
# template_raster_1km_GDA[template_raster_1km_GDA > 0] <- 1
# template_raster_1km_GDA[template_raster_1km_GDA < 0] <- 1
# xres(template_raster_1km_GDA)


```

\

Next we filter the records to those taken after 1950, and those inside the raster boundaries 
(i.e. species records in the ocean according to the raster boundaries will be excluded).

\

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=FALSE}

## Combine ALA data, and filter to records on land taken > 1950
## The climate data is the worldclim version 2.0
## Also, add a switch for the 1km filter - we don't want to do this anymore, too few records
data('ALA_keep')


## Combine Family ALA data
ALA.LAND.FAM <- combine_ala_records(species_list      = analysis_taxa,
                                    records_path      = "./data/ALA/Insects/Family/",
                                    records_extension = "_ALA_records.RData",
                                    record_type       = "ALA",
                                    keep_cols         = ALA_keep,
                                    world_raster      = world_annual_temp)


## Combine Genus ALA data
ALA.LAND.GEN <- combine_ala_records(species_list      = analysis_taxa,
                                    records_path      = "./data/ALA/Insects/Genus/",
                                    records_extension = "_ALA_records.RData",
                                    record_type       = "ALA",
                                    keep_cols         = ALA_keep,
                                    world_raster      = world_annual_temp)


## Combine Species ALA data
ALA.LAND.SPP <- combine_ala_records(species_list      = analysis_taxa,
                                    records_path      = "./data/ALA/Insects/Species/",
                                    records_extension = "_ALA_records.RData",
                                    record_type       = "ALA",
                                    keep_cols         = ALA_keep,
                                    world_raster      = world_annual_temp)

``` 

\

# STEP 3 :: extract environmental values

\

Next we combine occurrence files from ALA and GBIF into one table, and extracts environmental values. It assumes that both files come from the combine_ala_records and combine_gbif_records functions. Note that the order of the raster names in 'world_raster' must match the order of names in the character vector 'env_variables'. In this case, it's simply the biolclim variables (i.e. bio1-bio19)

\

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=FALSE}

## Combine GBIF and ALA data, and extract environmental values
## Note that the climate values are too small
COMBO.RASTER.FAM = combine_records_extract(ala_df          = ALA.LAND.FAM,
                                           site_df         = all_insect_site_df_families,
                                           thin_records    = FALSE,
                                           template_raster = template_raster_1km_WGS84,
                                           world_raster    = aus.climate.veg.grids,
                                           prj             = CRS("+init=epsg:4326"),
                                           species_list    = analysis_taxa,
                                           
                                           ## This might need to change too
                                           raster_divide    = FALSE,
                                           save_data        = FALSE,
                                           save_run         = "TARGET_INSECT_FAMILIES",
                                           data_path        = "./output/results/")


COMBO.RASTER.GEN = combine_records_extract(ala_df          = ALA.LAND.GEN,
                                           site_df         = all_insect_site_df_genera,
                                           thin_records    = FALSE,
                                           template_raster = template_raster_1km_WGS84,
                                           world_raster    = aus.climate.veg.grids,
                                           prj             = CRS("+init=epsg:4326"),
                                           species_list    = analysis_taxa,
                                           
                                           ## These two will need to change.
                                           ## Specify them in the code
                                           raster_divide    = FALSE,
                                           save_data        = FALSE,
                                           save_run         = "TARGET_INSECT_GENERA",
                                           data_path        = "./output/results/")


COMBO.RASTER.SPP = combine_records_extract(ala_df          = ALA.LAND.SPP,
                                           site_df         = all_insect_site_df_species,
                                           thin_records    = FALSE,
                                           template_raster = template_raster_1km_WGS84,
                                           world_raster    = aus.climate.veg.grids,
                                           prj             = CRS("+init=epsg:4326"),
                                           species_list    = analysis_taxa,
                                           
                                           ## These two will need to change.
                                           ## Specify them in the code
                                           raster_divide    = FALSE,
                                           save_data        = FALSE,
                                           save_run         = "TARGET_INSECT_SPECIES",
                                           data_path        = "./output/results/")

``` 

\

# STEP 4 :: Automated cleanin' of outlier records

\

The workflow uses four shapefiles as part of analysis and mapping: Australia, the World, 
the global Koppen Zones. The Koppen data are from CliMond, centred on 1975: 
https://www.climond.org/Core/Authenticated/KoppenGeiger.aspx

\

The next stage of the workflow use a series of cleaning functions to automate the removal 
of records for each species which are outliers. Doing this manually is extremely tedious, 
and although errors will be made, automation is preferable across large suites of taxa. 
The first cleaning function takes a data frame of all species records, and flag records 
as institutional or spatial outliers. This function uses the CoordinateCleaner package: 
https://cran.r-project.org/web/packages/CoordinateCleaner/index.html. It takes the 
records data.frame is that returned by the combine_records_extract function above.

\

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=FALSE}

## Just create a background points version for each step 

## Flag records as institutional or spatial outliers
COORD.CLEAN = coord_clean_records(records    = COMBO.RASTER.SPP,
                                  capitals   = 10000,  
                                  centroids  = 5000,   
                                  save_data  = FALSE,
                                  save_run   = "TARGET_INSECT_SPECIES",
                                  data_path  = "./output/results/")

``` 

\

The next cleaning function takes a data frame of all species records, flags records as 
spatial outliers (T/F for each record in the df), and saves images of the checks for each. 
Manual cleaning of spatial outliers is very tedious, but automated cleaning makes mistakes, 
so checking is handy. This function uses the CoordinateCleaner package https://cran.r-project.org/web/packages/CoordinateCleaner/index.html. 
It assumes that the input dfs are those returned by the coord_clean_records function.

\

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=FALSE}

## Flag spatial outliers
## This one doesn't quite work - it doesn't loop over the figures :
SPATIAL.CLEAN = check_spatial_outliers(all_df       = COORD.CLEAN,
                                       land_shp     = LAND,
                                       site_df      = FALSE, 
                                       clean_path   = './data/ALA/Check_Plots/',
                                       plot_points  = FALSE,
                                       record_limit = 300000,
                                       spatial_mult = 10,
                                       prj          = CRS("+init=epsg:4326"))

``` 

\

The next cleaning function takes a data frame of all species records, estimates the geographic 
and environmental ranges for each species, and creates a table of all species ranges. 
It uses the AOO.computing function in the ConR package: https://cran.r-project.org/web/packages/ConR/index.html
It assumes that the input df is that returned by the check_spatial_outliers function.

\

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=FALSE}

## Estimate climate niches using species records
## Make this calculate niches for only the target list
GLOB.NICHE = calc_1km_niches(coord_df     = SPATIAL.CLEAN,
                             prj          = CRS("+init=epsg:4326"),
                             country_shp  = AUS,
                             world_shp    = LAND,
                             kop_shp      = Koppen_shp,
                             species_list = analysis_taxa,
                             env_vars     = env_variables,
                             cell_size    = 2,
                             save_data    = TRUE,
                             save_run     = "ALL_INSECT_GENERA",
                             data_path    = "./output/results/")


## Blend the niches together
# GLOBAL_NICHES_1 <- readRDS('./output/results/GLOBAL_NICHES_ALL_INSECT_SPECIES.rds')
# GLOBAL_NICHES_2 <- readRDS('./output/results/GLOBAL_NICHES_ALL_INSECT_GENERA.rds')
# GLOBAL_NICHES_3 <- readRDS('./output/results/GLOBAL_NICHES_ALL_INSECT_FAMILIES.rds')
# GLOBAL_NICHES   <- bind_rows(GLOBAL_NICHES_1, 
#                              GLOBAL_NICHES_2,
#                              GLOBAL_NICHES_3) %>% .[.$searchTaxon %in% analysis_taxa, ]
# saveRDS(GLOBAL_NICHES, './output/results/ALL_INSECT_NICHES.rds')



``` 

\

We can also plot the environmental ranges of each species. This function 
takes a data frame of all species records, and plots histograms and convex hulls for 
each species in global environmental space. It assumes that the input df is that 
prepared by the check_spatial_outliers function

\

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=FALSE}

## Plot species ranges using histograms and convex hulls for rainfall and temperature distributions
plot_range_histograms(coord_df     = SPATIAL.CLEAN,
                      species_list = analysis_taxa,
                      range_path   = check_dir)
``` 

\


# STEP 5 :: Prepare SDM table

\

Then we need to create at table we can use for species distribution modelling. This function takes 
a data frame of all species records, and prepares a table in the 'species with data' (swd) format for 
modelling uses the Maxent algorithm. It assumes that the input df is that returned by the 
coord_clean_records function.

\

```{r message=TRUE, echo=TRUE, warning=FALSE, eval=FALSE}

## Two projection systems 
sp_epsg54009 <- "+proj=moll +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs +towgs84=0,0,0"
sp_epsg3577  <- "+proj=aea +lat_0=0 +lon_0=132 +lat_1=-18 +lat_2=-36 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs"

## Final dataset is a spatial points data frame in GDA Albers projection
SDM.SPAT.OCC.BG = prepare_sdm_table(coord_df        = COORD.CLEAN,
                                    species_list    = unique(COORD.CLEAN$searchTaxon),
                                    sdm_table_vars  = sdm_table_vars <- c('searchTaxon', 'lon', 'lat', 'SOURCE',
                                                                          names(aus.climate.veg.grids)),
                                    save_run        = "TARGET_INSECT_SPP",
                                    read_background = FALSE,
                                    save_data       = TRUE,
                                    save_shp        = FALSE,
                                    data_path       = paste0(getwd(), '/output/results/'),
                                    project_path    = 'G:/North_east_NSW_fire_recovery')

## Save shape file separately - doesn't seem to work in R
SDM.SPAT.OCC.BG.GDA <- spTransform(SDM.SPAT.OCC.BG, crs(sp_epsg3577))
writeOGR(obj    = SDM.SPAT.OCC.BG.GDA,
         dsn    = 'G:/North_east_NSW_fire_recovery/output/results',
         layer  = 'SDM_SPAT_OCC_BG_GDA_ALL_SPP', 
         driver = 'ESRI Shapefile', overwrite_layer = TRUE)

``` 

\


